BVP Paper

### 2017-11-27 tex/bvp_paper.tex
placeholders Jeff as an author
work on experiment section

\begin{document}
\author{Evan H. Anders}
\affiliation{Dept. Astrophysical \& Planetary Sciences, University of Colorado -- Boulder, Boulder, CO 80309, USA}
\affiliation{Laboratory for Atmospheric and Space Physics, Boulder, CO 80303, USA}
\author{Benjamin P. Brown}
\affiliation{Dept. Astrophysical \& Planetary Sciences, University of Colorado -- Boulder, Boulder, CO 80309, USA}
\affiliation{Laboratory for Atmospheric and Space Physics, Boulder, CO 80303, USA}
\title{BVP Paper}

@@

\section{Experiment} 
\label{sec:experiment}
Here we talk about how to set up the RB problem that we're going to solve, and exactly what
we're going to compare (how long the run is going to go from base and from bvp, how much time
we're going to compare over after that, etc.)  We need to compare over a short enough timescale
that the effects of BVPs don't get washed away, and we also need to compare over long enough
timescales that we get the rare events we expect.  This is where we put a figure showing the
time evolution of energy in IVP / BVP side-by-side.

>>>>

\begin{document}
\author{Evan H. Anders}
\affiliation{Dept. Astrophysical \& Planetary Sciences, University of Colorado -- Boulder, Boulder, CO 80309, USA}
\affiliation{Laboratory for Atmospheric and Space Physics, Boulder, CO 80303, USA}
\author{Benjamin P. Brown}
\affiliation{Dept. Astrophysical \& Planetary Sciences, University of Colorado -- Boulder, Boulder, CO 80309, USA}
\affiliation{Laboratory for Atmospheric and Space Physics, Boulder, CO 80303, USA}
\author{Jeffrey Oishi}
\affiliation{Bates}
\title{BVP Paper}

@@

\section{Experiment} 
\label{sec:experiment}
\subsection{Incompressible, Boussinesq \RB Convection}
In studies of convection, it is natural to nondimensionalize the flows on the freefall velocity.
Under the Boussinesq approximation, where $\rho = \rho_0 (1 - \alpha T_1)$ on the gravitational term in the momentum
equation and $\rho = \rho_0$ on all other terms, the freefall velocity is $v_{\text{ff}} = \sqrt{\alpha g L^2 (dT/dz)_0}$.
where $L$ is the depth of the domain, $g$ is the gravitational acceleration, and $(dT/dz)_0$ is the initial temperature
gradient of the atmosphere, which controls the flux through the system.  The typical timescale and length scale
then relate to each other according to $t_{\text{ff}} = L/v_{\text{ff}}$.
Under this nondimensionalization, the Boussinsq, incompressible equations of motion are
\begin{gather}
\Div{\bm{u}} = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\grad\varpi + T_1\hat{z} + \frac{\text{Pr}}{\text{Ra}}\grad^2\bm{u}, 
	\label{eqn:bouss_momentum_eqn}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \frac{1}{\text{Pr}\,\text{Ra}}\grad^2 T_1,
	\label{eqn:bouss_energy_eqn}
\end{gather}
where
\begin{equation}
\text{Ra} = \frac{g \alpha L^4 \left(\frac{dT}{dz}\right)_0}{\nu\chi} = \frac{(L\,v_{\text{ff}})^2}{\nu\chi}, \qquad \text{Pr} = \frac{\nu}{\chi}
\end{equation}
are the nondimensional Rayleigh and Prandtl numbers, where $\nu$ is the viscous diffusivity and $\chi$ is
the thermal diffusivity.

One hallmark of Boussinesq convection is that upflows and downflows are symmetric, and thus the long time- and horizontal-
average of the velocity field is exactly zero.  Assuming that the velocity field has perfect symmetry over a long time
average, the horizontally-averaged, steady state momentum equation is just
\begin{equation}
\frac{\partial}{\partial z}\angles{\varpi} = \angles{T_1}\hat{z},
	\label{eqn:simple_momentum}
\end{equation}
where $\angles{A}$ represent the horizontal- and time- average of the quantity $A$.  Thus, in the evolved solution, the
vertical profile of the pressure perturbations are directly determined by the temperature field. 
Under the Boussinesq approximation, $\bm{u}\cdot\grad(A) = \Div{\bm{u}A}$, and thus the steady state energy equation
is just
\begin{equation}
\frac{\partial}{\partial z}\angles{w(T_0 + T_1)} = \frac{1}{\text{Pr}\text{Ra}}\frac{\partial^2}{\partial z^2} \angles{T_1}.
	\label{eqn:simple_energy}
\end{equation}
Eqns (\ref{eqn:simple_momentum}) \& (\ref{eqn:simple_energy}) combine to form a second-order ODE which specifies the
mean vertical profile of the temperature and pressure fields in the evolved solution.  These fields are completely specified
by a set of two thermal boundary conditions and the evolved 
profile of the convective flux, $F_{\text{conv}} = \angles{w(T_0 + T_1)}$.

Under the choice of mixed thermal boundary conditions ($dT_1/dz = 0$ at $z = -1/2$, $T_1 = 0$ at $z = +1/2$), the initial
atmosphere starts off with much more thermal energy ($\propto T$) than the evolved adiabatic solution pegged by the
top temperature boundary condition.  This excess energy must leave the system through the upper (fixed $T$) boundary
during the convective transient, and this results in an asymmetric flux profile during the slow thermal evolution of the
atmosphere.  Furthermore, the convective flux (and the flux at the upper boundary layer) are O(1) during the convective
transient, whereas the flux entering the atmosphere at the lower boundary is O(Ra$^{-1/2}$), so the asymmetry of the fluxes
becomes increasingly pronounced as Ra is increased into the turbulent realm.

In order to find the evolved temperature profile of the atmosphere, the evolved profile of the convective flux must
be known.  We know that the evolved solution should be in flux equilibrium, and we know that the evolved flux through
the atmosphere is the flux entering through the bottom.  Thus, the steady-state profile of the convective flux can be
approximated as
\begin{equation}
F_{\text{conv, steady}} = F_{\text{bot}}\frac{\angles{w(T_0 + T_1)}}{\angles{w(T_0 + T_1) - \kappa \partial_z (T_0 + T_1)}}
= F_{\text{bot}}\frac{\angles{F_{\text{conv, IVP}}}}{\angles{F_{\text{tot, IVP}}}}.
\end{equation}
Or, put simply, the steady state convective flux is what you get get rid of the asymmetry in the flux profile.
This flux defined here is used in a 1D boundary value problem solve of 
eqns (\ref{eqn:simple_momentum}) \& (\ref{eqn:simple_energy}) along with the appropriate boundary values.  The solution
to that BVP is then used as new initial conditions, convection restarts, and so on.
